Improving \textsl{Gaia} parallaxes with a data-driven model of the color--magnitude diagram

### 2017-03-14 paperDistance/paper.tex

The novel use of \xd\ here is to treat its output---the deconvolved \cmd---as
a prior for use in inference of individual-star distances.
These inferences, one per star, provide much more precise parallax, distance,
or absolute magnitude estimates than we get from each star's primary
\tgas\ parallax measurement alone.
That is, we are using the \xd\ model to de-noise the \gaia\ parallax
measurements.
Technically, since \xd\ produces a maximum-marginalized-likelihood estimator
of the deconvolved distribution,
its use as a prior is only approximately valid; it constitutes using the
data twice.
However, since the data set is large, the \xd\ approximation is not bad; it is
sometimes known as the ``empirical Bayes'' method, and is well studied (DWH REFS).

@@

where $E(B-V)$ is the output of the Schlafly dust model, and $Q$ is the correction factor for that band assuming some reddening law. To generate our prior, we fit the data in the CMD space using XD (Bovy et al) which deconvolves the data with its noise, which are heterogeneous measurement uncertainties. It generates the denoised, underlying distribution from which the noisey data were drawn from. It represents the distribution as a mixture of gaussians which maximizes the likelihood of the noisy data.

>>>>

The novel use of \xd\ \citep{bovy11} here is to treat its output---the deconvolved \cmd---as
a prior for use in inference of individual-star distances.
These inferences, one per star, provide much more precise parallax, distance,
or absolute magnitude estimates than we get from each star's primary
\tgas\ parallax measurement alone.
That is, we are using the \xd\ model to de-noise the \gaia\ parallax
measurements.
Technically, since \xd\ produces a maximum-marginalized-likelihood estimator
of the deconvolved distribution,
its use as a prior is only approximately valid; it constitutes using the
data twice.
However, since the data set is large, the \xd\ approximation is not bad; it is
sometimes known as the ``empirical Bayes'' method, and is well studied (DWH REFS).

@@

where $E(B-V)$ is the output of the Schlafly dust model, and $Q$ is the correction factor for that band assuming some reddening law. To generate our prior, we fit the data in the CMD space using XD which deconvolves the data with its noise, which are heterogeneous measurement uncertainties. It generates the denoised, underlying distribution from which the noisey data were drawn from. It represents the distribution as a mixture of gaussians which maximizes the likelihood of the noisy data.
